Non-Vertical Cultural Transmission, Assortment, \\and the Evolution of Cooperation

### 2020-09-20 ms/main.tex
Update the figure and fix merge issues
Add condition for result 3 - when v<v_hat
Add alpha vs v to result figures to the notebook

Much of the literature on evolution of cooperation focuses on conditions for  initially rare cooperative phenotype to invade a population of defectors.
The next corollary deals with such a condition.
\\

\begin{corollary}[Condition for cooperation to increase when initially rare]
If the initial frequency of cooperation is very close to zero, $\tilde{p}_0 \approx 0$, then its frequency will increase if 
\begin{equation} \label{eq:unequal_transmission_from_rarity}
\begin{aligned}
c < \gamma_1 = \frac{b \alpha T_A + (T_A - T_B)}{1-T_B} \;.
\end{aligned}
\end{equation} 
\end{corollary}

Note that the right-hand side equals $\gamma_1$.
This condition follows from combining the conditions for fixation of cooperation and for stable coexistence.

In general, these conditions cannot be formulated in the form of Hamilton's rule due to the horizontal transmission bias $T_A-T_B$.
Without horizontal transmission bias, i.e., with $T=T_A=T_B$, these conditions reduce to a form of Hamilton's rule.
\\

@@

% in gamma
\begin{enumerate}
\item \emph{Fixation of cooperation}: 
	if \emph{(i)} $T_A \ge T_B$ and $c < \hat\gamma_1$; or 
	if \emph{(ii)} $T_A < T_B$ and $v>\hat v$ and $c < \hat\gamma_2$.
\item \emph{Fixation of defection}: 
    if \emph{(iii)} $T_A \ge T_B$ and $\hat\gamma_2 < c$; or 
	if \emph{(iv)} $T_A < T_B$ and $\hat\gamma_1 < c$.
\item \emph{Stable Coexistence}: 
    if \emph{(v)} $T_A < T_B$ and $v>\hat{v}$ and $\hat\gamma_2 < c < \hat\gamma_1$. 
\item \emph{Unstable coexistence}:
    if \emph{(vi)} $T_A > T_B$ and $\hat\gamma_1 < c < \hat\gamma_2$.
\end{enumerate}

@@

\emph{//additional corollaries//}

@@

\begin{figure}[htb]
  \centering       
    \includegraphics[width=0.45\textwidth]{Result3_TA_TB.eps}~
    \includegraphics[width=0.45\textwidth]{Result3_TB_TA.eps}
    \caption{\textbf{Evolution of cooperation under vertical, oblique, and horizontal cultural transmission.} 
The figure shows the global fixation of cooperation (green), global fixation of defection (red), fixation of either cooperation or defection depending on the initial conditions, i.e. unstable coexistence (blue), and stable coexistence of cooperation and defection (yellow).
The vertical transmission rate $v$ is on the x-axis.
Horizontal transmission bias (\textbf{a}) for cooperation, $T_A>T_B$ and (\textbf{b}) for defection, $T_A<T_B$.
Cost boundaries $\hat\gamma_1$ and $\hat\gamma_2$ (\autoref{eq:cost_boundaries_v}) are represented by the solid and dashed lines. 
(\textbf{a}) $b=1.2$, $T_A = 0.5$, $T_B = 0.4$, $\alpha = 0.4$.
(\textbf{b}), $b=2$, $T_A=0.5$, $T_B = 0.7$, $\alpha = 0.7$.
	}
    \label{fig:result3}
\end{figure}

>>>>

@@

% in gamma
\begin{enumerate}
\item \emph{Fixation of cooperation}: 
	if \emph{(i)} $T_A \ge T_B$ and $c < \hat\gamma_1$; or 
	if \emph{(ii)} $T_A < T_B$ and $v>\hat v$ and $c < \hat\gamma_2$.
\item \emph{Fixation of defection}: 
    if \emph{(iii)} $T_A \ge T_B$ and $\hat\gamma_2 < c$; or 
	if \emph{(iv)} $T_A < T_B$ and $\hat\gamma_1 < c$.
\item \emph{Stable Coexistence}: 
    if \emph{(v)} $T_A < T_B$ and $v>\hat{v}$ and $\hat\gamma_2 < c < \hat\gamma_1$; or
    if \emph{(vi)} $T_A < T_B$ and $v<\hat{v}$ and $c < \hat\gamma_1$.
\item \emph{Unstable coexistence}:
    if \emph{(vii)} $T_A > T_B$ and $\hat\gamma_1 < c < \hat\gamma_2$.
\end{enumerate}

@@

Much of the literature on evolution of cooperation focuses on conditions for  initially rare cooperative phenotype to invade a population of defectors.
The next corollary deals with such a condition.
\begin{corollary}[Condition for cooperation to increase when initially rare]
  If the initial frequency of cooperation is very close to zero, $\hat{p}_0 \approx 0$, then its frequency will increase if 
  \begin{equation} \label{eq:unequal_transmission_from_rarity_general_case}
  \begin{aligned}
  c < \hat\gamma_1 = \frac{b v \alpha T_A + (T_A - T_B)}{v(1-T_B)} \;.
  \end{aligned}
  \end{equation} 
  \end{corollary}

Note that the right-hand side equals $\hat\gamma_1$.
This condition follows from combining the conditions for fixation of cooperation and for stable coexistence.

In general, these conditions cannot be formulated in the form of Hamilton's rule due to the horizontal transmission bias $T_A-T_B$.
Without horizontal transmission bias, i.e., with $T=T_A=T_B$, these conditions reduce to a form of Hamilton's rule.
\\
\begin{corollary}[Necessity condition for fixation of cooperation.]
  Fixation of cooperation is possible only when
  \begin{equation} \label{eq:fixation_of_cooperation_vertical_transmission_condition}
    \begin{aligned}
      v>\hat{v} = \frac{T_B - T_A}{1-T_A} \;.
    \end{aligned}
    \end{equation} 
\end{corollary}
When there is horizontal transmission bias for cooperation, fixation can occured for every value of vertical transmission because $\hat{v}<0$.
However, if the the horizontal bias favours defection, for fixation of cooperation to occur the vertical transmission must be higer than $\hat{v}$.

@@

\begin{figure}[htb]
  \centering       
    \includegraphics[width=0.45\textwidth]{Result3_TA_TB.eps}~
    \includegraphics[width=0.45\textwidth]{Result3_TB_TA.eps}
    \caption{\textbf{Evolution of cooperation under vertical, oblique, and horizontal cultural transmission.} 
    The figure shows the global fixation of cooperation (green), global fixation of defection (red), fixation of either cooperation or defection depending on the initial conditions, i.e. unstable coexistence (blue), and stable coexistence of cooperation and defection (yellow).
    The vertical transmission rate $v$ is on the x-axis.
    Horizontal transmission bias (\textbf{a}) for cooperation, $T_A>T_B$ and (\textbf{b}) for defection, $T_A<T_B$.
    Cost boundaries $\hat\gamma_1$ and $\hat\gamma_2$ (\autoref{eq:cost_boundaries_v}) are represented by the solid and dashed lines. 
    (\textbf{a}) $b=1.2$, $T_A = 0.5$, $T_B = 0.4$, $\alpha = 0.4$.
    (\textbf{b}), $b=2$, $T_A=0.5$, $T_B = 0.7$, $\alpha = 0.7$.}
    \label{fig:result3}
\end{figure}
